---
title: "ECON 329 R Bootcamp Solutions"
author: "Matt Kaye"
date: "4/3/2019"
output: pdf_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

#put message = F and warning = F in this chunk
#so that you don't print warning or package startup messages
```

```{r install.packages, eval = FALSE}
check.packages <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
#install if you don't have these
#this chunk is only here for install help -- don't put install chunks in your RMD
#use the console to install the packages instead
check.packages("tidyverse")
check.packages("gapminder")
check.packages("stargazer")
check.packages("kableExtra")
```

#This is how you make a header -- more \# signs will make the header smaller
```{r load.packages}
#empty working environment to keep R workspace clean
rm(list = ls())

#load packages
library(gapminder)
library(tidyverse)
library(stargazer)
library(knitr)

#sets seed for random number generator so your bootstraps, random samples, etc
#will always knit the same
set.seed(10065) 
```

#Question 2 and Inline R!
```{r summary.stats, results='asis'}
#need results = 'asis' for stargazer(type = "latex") to work

df <- gapminder

df %>% #pipe operator -- makes df (left side) the first parameter in data.frame()
  data.frame() %>%
  select(4:6) %>% #use select() from dplyr to pick columns 4 through 6
  stargazer(header = FALSE, font.size = 'footnotesize') #stargazer summary stats table
```

You should type your answers to questions here, like this. If you have something in an R chunk that you want to print out as part of an answer (like the mean of gdpPercap, for example), you can print it inline like this: $`r round(mean(df$gdpPercap), digits = 3)`$. This will come in especially handy if you mess something up in question 1 that causes all of your answers to later questions to be off... Instead of needing to change everything, inline R will fix update everything for you.

#Question 3

```{r q1, fig.align="center"}
df %>% #use the pipe operator -- passes df into filter as first parameter
  filter(year == 2007) %>% #subset data to only 2007
  ggplot(aes(x = lifeExp, y = gdpPercap))+ #create ggplot
  geom_point() #add layer for point (i.e. scatter plot)
```
\newpage

#Question 4

```{r q2, fig.align = "center"}
df %>%
  #use %in% with filter to keep multiple years
  filter(year %in% c(1957, 1977, 1997, 2007)) %>% 
  #add colors and sizes
  ggplot(aes(x = lifeExp, y = gdpPercap, color = continent, size = pop))+ 
   #alpha blend the points -- make them translucent to make the plot clearer
  geom_point(alpha = .5)+
  scale_y_log10()+ #put the y on log scale
  #facet wrapper -- makes one plot for each category of the variable after "~"
  facet_wrap(~year, ncol = 2, nrow = 2)+ 
  #add plot labels
  labs(x = "Life Expectancy", 
       y = "GDP Per Capita (log scale)", 
       title = "GDP vs. Life Expectancy")
```

\newpage 

#Question 5
```{r, fig.align = "center"}
df %>%
  filter(year %in% c(1957, 2007)) %>%
  group_by(continent, year) %>% #sets continent and year as "grouping variables"
                                #lets us perform operations on every row within each group
  summarise(gdp = mean(gdpPercap)) %>% #calc mean gdpPercap for each continent-year pair
  spread(year, gdp) %>%  #create two new columns from the data frame we created above
  kable(format = 'latex', digits = 0) %>% #use kable to create nicely formatted tables
  kableExtra::row_spec(0, bold = TRUE) %>% #made header bold
  kableExtra::kable_styling(latex_options = "striped") %>% #made table striped
  kableExtra::add_header_above("GDP Per Capita")
```
\newpage

#Question 6:
```{r, fig.align="center"}
codes <- country_codes
df <- df %>%
  #left join attaches all rows matching the "by" in table 2 to table 1
  left_join(codes, by = "country") %>% 
  select(2, 1, 7, 3:6) %>% #selects columns by index
  mutate(pop = pop / 1000) #adds column

df %>%
  filter(year == 2007) %>%
  sample_n(10, FALSE) %>% #random sample of 10 rows without replacement
  arrange(continent, country) %>%
  kable(format = 'latex', digits = 0) %>% #use kable to create nicely formatted tables
  kableExtra::row_spec(0, bold = TRUE) %>% #made header bold
  kableExtra::kable_styling(latex_options = "striped") %>% #made table striped
  kableExtra::add_header_above("Random Country Sample") %>% #add title
  kableExtra::footnote(symbol = "population in thousands") #add a note
```


#Appendix Of Useful Latex Stuff


If you want to make some text \textbf{bold} or \textit{italic}, do this

If you want to make your regression equations look nice, do something like this:

$$Y = \hat{\beta_{0}} + \hat{\beta_{1}}\cdot X + \epsilon$$

The dollar signs tell Rmarkdown to use LaTeX's math mode, which lets you use Greek letters, subscripts, etc.

\indent lets you indent a line
\noindent stops a line from being indented.

\begin{itemize}
\item lets you add bullet points
\end{itemize}

\begin{enumerate}
\item same idea for adding numbered bullets
\end{enumerate}

if you want to make equations really fancy (to do derivations of properties of summations, for example), you can do something like this:

\begin{center}
\begin{align*}
X &= Y\\
Y &= Z\\
\implies X &= Z
\end{align*}
\end{center}

\begin{center}
this is how you center text
\end{center}

and this is how you skip some space

\vspace{1cm}

Now, some potentially helpful side notes: First, Detexify (for LaTeX) and StackOverflow are great resources for help on LaTeX and R questions. Also, the R documentation on all of these functions is really good. Doing something line ?filter or ?ggplot in the R console will open the documentation (or you can just Google it).






